---
title: "Prediction of Human Activity Recognition (KNN)"
author: "Qing (Richard) Deng"
date: "November 12, 2015"
output: html_document
---

Practical Machine Learning Project

## Data source and downloading
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

```{r}
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
#download.file(url,"/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-training.csv")
train_pml <- read.csv("/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-training.csv")

url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url,"/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-testing.csv")
test_pml <- read.csv("/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-testing.csv")
```


## FEATURE SELECTION
First, we clean up missing values, remove zero covariate and descriptive features. 
```{r}
        # DEAL COMLUMNS WITH NAs
        na <- sapply(test_pml, function(x){sum(is.na(x))})
        training_knn <- subset(train_pml, select = !(names(train_pml) %in% names(na[na==20])))
        testing_knn <- subset(test_pml, select = !(names(test_pml) %in% names(na[na==20])))
        
        # REMOVE DESCREIPTIVE FEATURES. 
        excludecols <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
                         "cvtd_timestamp", "num_window", "new_window")
        training_knn <- training_knn[, !names(training_knn) %in% excludecols]
        testing_knn <- testing_knn[, !names(testing_knn) %in% excludecols]
        
        # NORMALIZE DATA both training_knn and testing_knn
        normalize <- function(x) { return ((x - min(x)) / (max(x) - min(x)))}
        tem <- rbind(training_knn[,-53], testing_knn[-53])
        t_n <- as.data.frame(lapply(tem, normalize))
        training_knn<- cbind(t_n[-20:-1,], training_knn[53])
        testing_knn <- t_n[19623:19642,]
        
        # DATA SPLITING
        library(caret); set.seed(1001)
        intrain <- createDataPartition(y=training_knn$classe, p=0.8, list = F)
        training_body <- training_knn[intrain,][,-53]
        validation_body <- training_knn[-intrain,][,-53]
        training_label <- training_knn[intrain,][,53]
        validation_label <- training_knn[-intrain,][,53]
```

# VALIDATION SET PREDICTION and EVALUATION
We will us KNN as our model IMPLEMENTED IN THE class PACKAGE
```{r}
library(class)
pred_vlidation<- knn(training_body,validation_body, training_label, k=80 )
print(confusionMatrix(pred_vlidation, validation_label))  
```

# PERFORMACE INCREASE

```{r}
    # try different K - k=20    
    pred_vlidation<- knn(training_body,validation_body, training_label, k=20 )
    print(confusionMatrix(pred_vlidation , validation_label))
```

```{r}
    # try different K - k=2 
    pred_vlidation<- knn(training_body,validation_body, training_label, k=5 )
    print(confusionMatrix(pred_vlidation , validation_label))  
```

# PERFORMANCE INCREASE 2
```{r}
    # USE Z-SCORE SDANDARDIZE DATA
    t_z <- as.data.frame(scale(training_knn[,-53]))
    training_knn_z<- cbind(t_z, training_knn[53])
    
    # DATA SPLITING
    library(caret); set.seed(1001)
    intrain <- createDataPartition(y=training_knn_z$classe, p=0.8, list = F)
    training_body <- training_knn_z[intrain,][,-53]
    validation_body <- training_knn_z[-intrain,][,-53]
    training_label <- training_knn_z[intrain,][,53]
    validation_label <- training_knn_z[-intrain,][,53]  
    
    # PREDICTING
    # when k=50
    pred_vlidation<- knn(training_body,validation_body, training_label, k=80 )
    print(confusionMatrix(pred_vlidation , validation_label))  
    
    # when k=20
    pred_vlidation<- knn(training_body,validation_body, training_label, k=20 )
    print(confusionMatrix(pred_vlidation , validation_label))  
    
    # when k=5
    pred_vlidation<- knn(training_body,validation_body, training_label, k=5 )
    print(confusionMatrix(pred_vlidation , validation_label))
```