---
title: "Prediction of Human Activity Recognition (Random Forest)"
author: "Qing (Richard) Deng"
date: "November 12, 2015"
output: html_document
---
Practical Machine Learning Project

# INTRODUCTION

Using fitness wearable devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively. These type of can help people to find patterns in their behavior, quantifying how much of a particular activity they do. But they rarely quantify how well they do it. 

In a research, six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different ways as follows:: 

Class A: exactly according to the specification.

Class B: throwing the elbows to the front.

Class C: lifting the dumbbell only halfway.

Class D: lowering the dumbbell only halfway.

Class E: throwing the hips to the front.

In this project, our goal is to to predict how well an activity was performed by the wearer. 

## Data source and downloading
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

```{r}
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
#download.file(url,"/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-training.csv")
train_pml <- read.csv("/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-training.csv")

url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url,"/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-testing.csv")
test_pml <- read.csv("/Users/Richard/Desktop/Rcourse/data/practical_machine_l/pml-testing.csv")
```

The training data has 19622 observations and 160 features, and the distribution of the five measured stances A,B,C,D,E is:
```{r}
dim(train_pml)
table(train_pml$classe)
```

# PREPROCESSING
## DATA SPLITING
We separate our training data into a training set and a validation set so that we can validate our model.
```{r}
    library(caret); set.seed(1001)
    intrain <- createDataPartition(y=train_pml$classe, p=0.8, list = F)
    training <- train_pml[intrain,]
    validation <- train_pml[-intrain,]
```

## FEATURE SELECTION
First, we clean up missing values, remove zero covariate and descriptive features. 
```{r}
# DEAL COMLUMNS WITH NAs
na <- sapply(test_pml, function(x){sum(is.na(x))})
training <- subset(training, select = !(names(training) %in% names(na[na==20])))
# REMOVE ZERO COVARIATE
nsv <- nearZeroVar(training)
training <- training[,-nsv]
# REMOVE DESCREIPTIVE FEATURES. 
excludecols <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
                  "cvtd_timestamp", "num_window")
training <- training[, !names(training) %in% excludecols]
```

Data structure looks like this:
```{r}
str(training)
```

# MODEL TRAIN(RANDOM FOREST) 
We will use random forest as our model implemented in the randomForest package
```{r}
library(randomForest)
rfModel <- randomForest(classe~., data=training, importance= T, ntrees=10)
```

# MODEL VALIDATION - VALIDATION SET ACCURACY
Let us now test our model performance on the validation set.
```{r}
pre_validation <- predict(rfModel,validation)
print(confusionMatrix(pre_validation, validation$classe)) ## acuurancy = 0.9967
```

# TEST SET PREDICTION
The prediction of our algorithm for the test set is:
```{r}
pre_test <- predict(rfModel,test_pml)
pre_test
```

# OUTPUT THE RESULT ACCORDING TO THE INSTRUCTURE. 
We then save the output to files according to instructions and post it to the submission page.
```{r}
answers <- as.vector(pre_test)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
                    col.names = FALSE)
    }
}

pml_write_files(answers)
```

```{r}

```

```{r}

```

```{r}

```

